Calculating the Solvent Exposed Surface Area of Proteins with Intersecting Spheres

### 2020-06-06 Paper.tex
More Edits
more changes

In a previous paper written by Richmond (1984), a different strategy is used to find the solvent exposed surface area of the atoms. He still represents the atoms as spheres with definite radius with centers at the atomic coordinates, but instead of using the Alpha Shapes, he made use of the Gauss-Bonnet theorem, a fundamental theorem in differential geometry which connects the topological properties of a manifold (specifically its Euler Characteristic) with its geometric properties (its curvature). A standard statement of the theorem is 

@@

The SASA of an atom with radius $\varrho_i$ lies on a sphere where $K=\frac{1}{\varrho_i^2}$. We can divide this SASA into many polygons bounded by the circles of intersection with neighboring spheres (See Figure). If $M$ is one particular region of that SASA, we can write 
\begin{align}\nonumber
&\frac{1}{\varrho_i^2}\int_M dA + \int_{\partial M} k_g ds= 2\pi \\
&\implies A=\varrho_i^2\left(2\pi-\int_{\partial M} k_g ds\right)
\end{align}

@@

\begin{figure}[h!]
\caption{The dual complex (K) of the Delaunay Triangulation only includes those power cells that non-empty intersection.$\lbrack1\rbrack$}
\centerline{\includegraphics[width=3in]{Figure9}}
\end{figure}

As can be seen by figure 6, K is made up of multiple simplices. The star of K is a set of triangles who share a vertex and outside boundary. The link of K, then, is all of the vertices and edges in the star (Figure 7).

\begin{figure}[h!]
\caption{The star of K and link of the star are shown.$\lbrack1\rbrack$}
\centerline{\includegraphics[width=4in]{Figure5}}
\end{figure}

@@

\begin{align}
\sigma_{ij} = 1 - \frac{\sum_{k}length(S^k_{ij})- \sum_{k,l}length(S^k_{ij} \cap S^l_{ij})}{2\pi\varrho_{ij}}\\
\end{align}

@@

\section{Glossary}
\begin{align*}
&\varrho_i:&\text{The radius of ball i} \\
&z_i:&\text{The center of a ball i} \\
&{\bf z}:&\text{The $3n$ dimensional vector representing the state of all $z_i$} \\
&\zeta_{ij}:&\text{The signed distance between centers of two intersecting spheres} \\
&\sigma_{ij}:&\text{The fraction of solvent-exposed surface for two intersecting spheres $S_i,S_j$} \\
&\beta_{ijk}:&\text{The fraction of solvent-exposed surface for three intersecting spheres $S_i,S_j, S_k$} \\
&\alpha:&\text{An atomic solvation parameter} \\
&F:&\text{The space filling diagram} \\
&\pi_i(x):&\text{The power distance of a random point x from $S_i$} \\
&P_i:&\text{The set of points closer (by power distance) to $S_i$ than any other sphere} \\
&K:&\text{The dual complex of Delaunay Triangulation. Consists of all power cells with non-empty intersection} \\
&DA_{\bf z}({\bf t}):&\text{The unweighted area derivative at \textbf{t}} \\
&DW_{\bf z}({\bf t}):&\text{The weighted area derivative at \textbf{t}} \\
\end{align*}

>>>>

In a previous paper written by Richmond (1984), a different strategy is used to find the solvent exposed surface area of the atoms. He still represents the atoms as spheres with definite radius with centers at the atomic coordinates, but instead of using the Alpha Shapes, he made use of the Gauss-Bonnet theorem, a fundamental theorem in differential geometry which connects the topological properties of a manifold (specifically its Euler Characteristic) with its geometric properties (its curvature) (Rotskoff 2010, Schnell 2004). A standard statement of the theorem is 

@@

The SASA of an atom with radius $\varrho_i$ lies on a sphere where $K=\frac{1}{\varrho_i^2}$. We can divide this SASA into many polygons bounded by the circles of intersection with neighboring spheres (See Figure 2). If $M$ is one particular region of that SASA, we can write 
\begin{align}\nonumber
&\frac{1}{\varrho_i^2}\int_M dA + \int_{\partial M} k_g ds= 2\pi \\
&\implies A=\varrho_i^2\left(2\pi-\int_{\partial M} k_g ds\right)
\end{align}

@@

\begin{figure}[h!]
\caption{The dual complex (K) of the Delaunay Triangulation only includes those power cells that have a non-empty intersection.$\lbrack1\rbrack$}
\centerline{\includegraphics[width=3in]{Figure9}}
\end{figure}

As can be seen by Figure 6, K is made up of multiple simplices. The star of K is a set of triangles who share a vertex and outside boundary. The link of K, then, is all of the vertices and edges in the star (Figure 7).

\begin{figure}[h!]
\caption{The star of K and link of the star are shown.$\lbrack1\rbrack$}
\centerline{\includegraphics[width=3in]{Figure5}}
\end{figure}

@@

\begin{equation*}
\sigma_{ij} = 1 - \frac{\sum_{k}length(S^k_{ij})- \sum_{k,l}length(S^k_{ij} \cap S^l_{ij})}{2\pi\varrho_{ij}}
\end{equation*}

@@

\section{Glossary}
\begin{align*}
&\varrho_i:&\text{The radius of ball i} \\
&z_i:&\text{The center of a ball i} \\
&{\bf z}:&\text{The $3n$ dimensional vector representing the state of all $z_i$} \\
&\zeta_{ij}:&\text{The signed distance between centers of two intersecting spheres} \\
&\sigma_{ij}:&\text{The fraction of solvent-exposed surface for two intersecting spheres $S_i,S_j$} \\
&\beta_{ijk}:&\text{The fraction of solvent-exposed surface for three intersecting spheres $S_i,S_j, S_k$} \\
&\alpha:&\text{An atomic solvation parameter} \\
&F:&\text{The space filling diagram} \\
&\pi_i(x):&\text{The power distance of a random point x from $S_i$} \\
&P_i:&\text{The set of points closer (by power distance) to $S_i$ than any other sphere} \\
&K:&\text{The dual complex of Delaunay Triangulation. Consists of all power cells with non-empty intersection} \\
&DA_{\bf z}({\bf t}):&\text{The unweighted area derivative at \textbf{t}} \\
&DE_{\bf z}({\bf t}):&\text{The weighted area derivative at \textbf{t}} \\
\end{align*}
